Scaling of Electrode-Electrolyte Interface Model Parameters In Phosphate Buffered Saline

### 2014-04-16 paper.tex
Another round of edits

\usepackage{cite}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{url}
\usepackage{ulem}
\usepackage[usenames, dvipsnames]{color}

@@

Electrodes in the laboratory are typically tested in a saline solution selected to mimic the circumstances in which they will operate when implanted.

@@

In an implanted setting there is no reference electrode and many of the well established electrochemical measurement regimes do not apply. In this manuscript we measure the behaviour an electrode/electrolyte situation as it would be seen from an implant device.

@@

{
    \color{blue}
\section{Compact Non-Linear Electrical Model}
\label{sect:resistorNetwork}
\begin{figure}
    \begin{center}
        \includegraphics[width=230pt]{graphics/interfaceSchematic_noMemristive}
    \end{center}
    \caption{Electrical schematic of the electrode-interface impedance model used in this work. $D_{a}$ and $D_{b}$ represent diodes; $CPE$ and $R_{s}$ represent the constant phase element and series resistance respectively.}
    \label{fig:schematic}
\end{figure}

In this manuscript we use an adapted version of the compact non-linear model presented by Scott and Single \cite{ScottSingle2013}. The model is compact in that it is suitable for entering into SPICE, i.e., freely available electrical circuit simulation software. The schematic for the model used here is presented be seen in Figure \ref{fig:schematic}.
}
The electrical interface model of \cite{Franks2005} and \cite{ScottSingle2013} has two parts, a displacement branch and a Faradaic branch.

Scott and Single \cite{ScottSingle2013} incorporate memristors into their model. {
    \color{blue}
    These memristors are used to mimic a diffusion limited current conduction mode, i.e., their purpose is to throttle electrical current associated with Faradaic reactions. In an implanted setting an electrode should never be operated in a Faradaic counduction mode and certainly never to the point of becoming diffusion limited; for this reason the memristors have been omitted.
}

% Contribution from memristive elements is limited to situations where diffusion limited Faradaic reactions come into play, a situation that stimulator implants are designed to avoid.{
%     \color{blue}
%     Modelling Faradaic current onset allows engineers to predict and avoid such situations in-vivo. For this reason the measurements and model fitting are primarily concerned with capturing the onset of Faradaic conduction, as is modelled by diodes. The model used here is not intended to model the behaviour of an electrodes operating in a Faradaic conduction mode. It is for this reason we have removed the memristors and limit the model to a single pair of diodes, thus capturing only the beginning of the primary Faradaic component.
% }

It has been suggested that the primary coupling mechanism between platinum electrodes and physiological saline is the reversible evolution of gas, probably oxygen, at the platinum surface, which dissolves without bubble formation.\cite{Greatbatch1969} Understanding the exact mechanism of displacement or Faradaic charge transfer at the electrode surface is beyond the scope of this work. Rather we wish to characterise the electrode/electrolyte system in terms of a {\color{blue} small number of} parameters that may be fed into an electrical model.

@@

The electrical impedance between two electrodes in an electrolyte arises from two interface impedances in series with the resistance due to the bulk of the electrolyte itself. To quantify the impedance presented by a single interface we first need to understand the inter-electrode resistances presented by the electrolyte bulk. As there are always two interface impedances between any two electrodes, inter-electrode resistances cannot be measured in isolation.

In \cite{ScottSingle2013}, a series of transresistance measurements were used in conjunction with a physical description of the electrode geometry to build up a representative network of resistors. The network in that paper was defined using five parameters: $R_{eri}$, $R_{lri}$, $R_{li}$, mesh depth, and edge depth. Readers are referred to \cite{ScottSingle2013} for the interpretation of these model parameters. Repeating that work, a resistor network that connects each of the eight electrodes is constructed, mimicking the resistances due to the solution bulk.

The mesh depth and padding values of five columns and three rows, respectively, has been taken directly from \cite{ScottSingle2013}. {\color{blue} Using these parameters together with an array of 8 electrodes leads to a resistor network having 205 resistors.}
A numerical fit was made to the two independent scaling parameters $R_{eri}$ and $R_{li}$
\footnote{The values of $R_{eri}$ and $R_{sri}$ are defined by the lengths of the electrode and inter-electrode spacers respectively as described in \cite{ScottSingle2013}. The two parameters are proportional to each other by the ratio of their lengths. Hence, we define $R_{sri}$ to be 3/4 that of $R_{eri}$ making it a dependant parameter.}.

The resulting fit and parameter values are presented in figure~\ref{fig:transresistance} and table~\ref{tab:RESparams} respectively.

Measurements of the inter-electrode impedance were made using a Tektronix TPS2014 DSO with four fully isolated and floating channels; an Agilent 33220A Function Generator; and a desktop PC controlling the instruments via Python 2.7.

The pH level of the stock solution was measured, using a calibrated EDT Instruments BA-350 pH meter, to have a pH of 7.4. No pH adjustments were made to the derived solutions.

    In Table \ref{tab:RESparams}, $\sigma$ refers to the conductivity of the solution, measured in units of Siemens per centimeter. Solution conductivity was measured using an EDT instruments RE 388 TX Microprocessor Conductivity Meter.

    It should be noted that the resistive mesh created in this step models only the resistivity of the solution bulk. This means that no resistive contribution from the electrode/electrolyte interface is included in these measurements as voltage measurements are always across pairs of non-driven electrodes. We determine the fixed resistance of a single interface, $R_{S}$, when fitting parameters to the displacement component.
}

@@

% \begin{table}
%     \caption{Measured conductivity of PBS solutions. {\color{red} Should I add this? The meter was uncalibrated}}
%     \label{tab:PBSconductivity}
%     \begin{center}
%         \begin{tabular}{r | l}
%             Concentration (X-PBS) & Conductivity ($\mu S/cm$)\\
%             \hline
%             1.0 & $15800$ \\
%             0.5 & $4360$\\
%             0.25 & $1860$\\
%             0.05 & $931$\\
%             0.025 & $492$\\
%         \end{tabular}
%     \end{center}
% \end{table}

@@

\section{Displacement Parameters {\color{blue}/ CPE}}
\label{sect:cpe}
110 elements
Displacement currents are responsible for capacitive behaviour at interfacial boundaries.
They are brought about by the redistribution of charge in response to applied fields, i.e. a charged electrode will attract or repel $Cl^{-}$ from its surface and reorientate polar molecules in the surrounding solution.\cite{Merrill05} This capacitive behaviour may also be a result of electrode polarisation in the form of reversible Faradaic reactions at the surface of the electrode.
Such reactions involving water and platinum, as identified in \cite{Horch2004,Mohtashami2011,Merrill05}, include:
\begin{eqnarray}
    Pt + H_{2}O &\Leftrightarrow& PtO + 2 H^{+} + 2 e^{-}\\
    PtO + H_{2}O &\Leftrightarrow& PtO_{2} + 2 H^{+} + 2e^{-}\\
    Pt + H^{+} & \Leftrightarrow & Pt-H\\
    Pt + H_{2}O + e^{-} &\Leftrightarrow& Pt-H+OH^{-}
\end{eqnarray}

This displacement mechanism is capable of transferring charge between electrical and biological systems without damaging electrodes or tissue.\cite{Horch2004}
This behaviour is modelled by way of a Constant Phase Element (CPE); also known as a fractional capacitor.
For simulation purposes this element is realised as an array of R-C branches.~\cite{ScottSingle2013,Morrison59,Elwakil10} where the value of each R-C pair is chosen to set the slope of impedance relative to frequency. The placement of R-C pairs within the CPE is determined by Morrison's parameters $m$ and $k$, where $k$ determines the spacing, or density, of R-C branches and $m$ determines the resulting slope. {\color{blue} The R-C pairs were generated via Python script for to span 1\thinspace nHz--1\thinspace GHz with a spacing of 3 pairs per decade. The final CPE contains 55 branches i.e. 110 elements are used to create the CPE model.}

@@

\begin{figure}
    \begin{center}
        \includegraphics{graphics/chargeInjectionVsFrequency_magnitude}
    \end{center}
    \caption{Charge injected per phase during CPE measurement for each of the six concentrations of PBS. Calculated using a geometric surface area of 10\thinspace mm$^{2}$}
    \label{fig:chargeInjectionVsFrequency}
\end{figure}

The geometric current density during measurements peaked at 22.6\thinspace mA/cm$^{2}$ at 10000\thinspace kHz in 1.0X PBS and fell as low as 27.6\thinspace $\mu$A/cm$^{2}$ at 50\thinspace mHz in 0.025X PBS. Charge injection per phase is shown in figure~\ref{fig:chargeInjectionVsFrequency} for each solution.
The stimulus waveform was set at 300\thinspace mV (peak) at each measurement point. The measurement instruments used are the same as those used to measure the inter-electrode resistance.
Figure~\ref{fig:CPE_Magnitude} shows the magnitude of the measured response of each of the six PBS concentrations accompanied by simulation results. Figure~\ref{fig:CPE_Phase} shows the phase response for the same measurements, again with the simulated response at each concentration. The simulated responses are generated using the optimised parameter values presented in Table~\ref{tab:CPEparams}. Those values were chosen as they give the closest match to the measured data.

@@

Our model uses two reverse connected diodes to represent Faradaic current conduction between electrode and electrolyte. Specifically we use the diodes to model irreversible Faradaic reactions. (Any reversible reactions, those producing reactants bound to the surface of the electrode, are encapsulated by the displacement branch (CPE) of the model.)
Possible irreversible Faradaic reactions for platinum electrodes in saline are:

@@

{
    \color{blue}
    It should be noted that each of these reactions (\ref{eqn:ptCl}, \ref{eqn:H20}, \ref{eqn:2H20}, \ref{eqn:Cl} and \ref{eqn:ClH20}) have different reaction potentials. It is expected that the situation in-vivo will be much more complex.
}

The interface model uses the diodes to reproduce the exponential behaviour of Faradaic current conduction implied by the Butler-Volmer equation. We wish to measure this conduction by separating it from the displacement behaviour of the CPE. By using a DC measurement technique, as opposed to cyclic measurements, we are able to separate the effects in time.
Like a conventional capacitor, when the voltage across the CPE changes, it responds by absorbing or ejecting charge, causing a spike in current. The CPE will draw negligible current once it has settled and in the case of our model we assume the remaining current conduction to be the result of Faradaic processes.
In \cite{Greatbatch1969} it is shown that the stirring of an air-saturated solution of physiological saline reduces the settling time of platinum electrodes to stimulus transients; with both situations eventually settling to the same impedance.
The equation governing current conduction in a diode where the forward voltage across the diode, $V_{D}$, is small is
\begin{equation}
    I = i_{0}  e^{V_{D} / n V_{T}}
\end{equation}
where $V_{T}$ is the thermal voltage and is approximately 25\thinspace mV at a temperature of 300\thinspace K. We wish to find how the saturation current ($i_{0}$) and the ideality factor ($n$) scale as the solution salinity is varied.

@@

@@

%Parameter values for $i_{0}$ and $n$ are presented in table~\ref{tab:FaradaicParams}. These parameters were found by fitting the diode equation to the results shown in figure~\ref{fig:faradaic_currentVsVoltage}.

@@

\begin{figure}
    \begin{center}
        \includegraphics{graphics/sheep_transimpedance_magnitude}
    \end{center}
    \caption{\color{blue} Comparison between the magnitude of in-vivo transresistance measurements (blue dots) and simulation results using fitted mesh parameters (red trace).}
    \label{fig:transimpedance_sheep_mag}
\end{figure}

\begin{figure}
    \begin{center}
        \includegraphics{graphics/sheep_transimpedance_phase}
    \end{center}
    \caption{\color{blue} Comparison between the phase of in-vivo transresistance measurements (blue dots) and simulation results using fitted mesh parameters (red trace). The simulation mesh does not capture any phase relationship due to being purely resistive.}
    \label{fig:transimpedance_sheep_phase}
\end{figure}

@@

{\color{blue}
    Four measurements of displacement current (CPE element in the model) were} taken at regular intervals over a period fo 30 minutes and have been presented in figures~\ref{fig:displacement_sheepCPEMagnitude} and~\ref{fig:displacement_sheepCPEPhase}.
{\color{blue} These measurements appear on top of the traces taken from figures~\ref{fig:CPE_Magnitude} and~\ref{fig:CPE_Phase} for comparison.}

@@

\begin{figure}
    \begin{center}
        \includegraphics{graphics/displacement-withSheep_impedanceVsFrequency_phase}
    \end{center}
    \caption{\color{blue} Comparison between the mean phase response of measured displacement currents in four live sheep (blue trace), each of the PBS traces taken from figure~\ref{fig:CPE_Phase} (grey lines), and a simulated fit to the sheep data (red trace). Error bars have been placed one standard deviation either side of the mean for measured data.}
    \label{fig:displacement_sheepCPEPhase}
\end{figure}

\section{Conclusion}
\label{sect:conclusion}
We have used the parameters of a compact electrical model of an implantable electrode as a way of objectively comparing scenarios.
We have measured how changing the concentration of PBS affects the parameters of this model.
We have drawn direct relationships with concentration {\color{blue} and conductivity} sufficient to predict interface characteristics at arbitrary dilutions of PBS.
We found that the magnitude of the CPE element moves much more slowly with concentration whereas the resistance moved almost linearly.
Using DC measurements we conclude that the {\color{blue} concentration of PBS has no effect on the onset of Faradaic conduction for the range of concentrations investigated.} {\color{Green} \sout{salinity of the bulk does not affect Faradaic conduction until a threshold voltage is reached}}. We hypothesise that this is because the double-layer sets the effective concentration of ions {\color{blue}, those involved in the Faradaic reactions,} in a localised volume surrounding each electrode.
In-vivo measurements of platinum electrodes in sheep spine show that no single concentration of PBS matches both the CPE and the resistive characteristics at once. {\color{blue} These measurements also reveal a considerable amount of reactance (see transimpedance measurements in-vivo) that is captured neither by PBS nor the model presented when using a purely resistive spreading resistance mesh.} Nevertheless, {\color{blue}if a single solution of PBS is to be used as a human phantom, a one-tenth concentration solution (by mass)} is a good compromise.

{
    \color{blue}
    The model is limited to predicting only the onset of Faradaic current conduction, i.e. it does not model operation in a diffussion limited scenario as it omits the memristors of \cite{ScottSingle2013}.
}

\section*{Acknowledgment}
The authors wish to acknowledge the The University of Waikato for financial support and Saluda Medical for the use of facilities and equipment. Thanks are extended to Peter Single for help conducting in-vivo measurements.

@@

\bibitem{ScottSingle2013}
{\color{red} HAS THIS BEEN PUBLISHED YET?}
    Jonathan Scott and Peter Single,
    ``Compact Nonlinear Model of an Implantable Electrode Array for Spinal Cord Stimulation (SCS)'',
    to be published in
    {\em IEEE Transactions on Biomedical Circuits and Systems},
    DOI: 10.1109/TBCAS.2013.2270179, 2013.

@@

    {
        \color{blue}
        The diodes are there to model the change in current attributed to the first biologically damaging reaction that occurs at the electrode/electrolyte interface. Whatever reaction occurs first in the relevant electrolyte is the reaction that diode aims to mimic. We observe that the current increases near exponentially around this point so we use a diode to mimic that. We expect this reaction may not be the same in-vivo as it is in-vivo but the properties of a diode can be altered to match where this occurs.

        If there are secondary reactions occurring at higher potentials then we could add a second pair of diodes with the properties of that specific reaction. The problem here is that we would then need to include memristors (as used in ScottSingle2013) to account for reactant depletion. We don't account for reactant depletion for the same reason we don't account for reactions other than the very first Faradaic reaction to occur. The reason is that we never want biologically damaging reactions to occur due to an implant, so if they happen then something has gone wrong.

        Think of our attempt to model Faradaic reactions as a ``you've pushed this too far'' indicator for an electrical engineer.

        I think you're essentially right with you're argument, we're not up-front about this and we probably portray more capability in the Faradaic modelling aspect than we are able or aim to achieve.
    }

@@

    {
        \color{blue}
        Correct, we do not wish to determine exactly what is happening at the electrode surface - we want to create a flexible model that can be used to simulate impedance.

@@

    {
        \color{blue}
        This is a good point. ScottSingle2013 compare the model prediction to stimulus waveforms. Our aim is to see what changes in that model as the concentration of PBS is varied and to compare those measurements to in-vivo measurements using a live sheep. Were doing this because 0.1X PBS is commonly used as a phantom for real spine. There hasn't been an objective way of determining how accurate that assumption was until we used this model to make that comparison.
    }

@@

    {
        \color{blue}
        We need this because it represents the resistance of the interface itself, separate from that of the fluid.
        We measure the impedance magnitude of the displacement currents at high frequency (flattened part of Fig. 5). For each concentration the impedance magnitude at 10kHz is higher than is predicted by the transimpedance measurements made in Fig. 3. The transimpedance measurements do not measure impedance across an interface that is passing current and therefore any voltage drop across an interface due to $R_{S}$ would vanish. The displacement measurements take the voltage across interfaces that are passing stimulus currents. These measurements see the potential drop across the interface due to the current flow via Ohms law.

@@

    {
        \color{red}
        What should I say to this?. We measured what we measured, no visible gas evolution. The results look consistant. We are happy to accept that we may have pushed the electrodes too far in the solution at low frequencies, regardless of this fact we have data here and wish to present it.
    }

@@

    {
        \color{blue}
        The RC pairs are generated via a Python script for a given frequency range. Simulations in this manuscript used RC pairs spanning 1.0 nHz to 1.0 GHz spaced 3 per decade. Adding more RC pairs than necessary came at negligible computation cost and having the circuited constructed in an automated fashion to satisfy the given frequency range adds no extra work.
    }

@@

   {
       \color{red}
       HELP, how do I approach this?
   }

@@

    9. In their conclusions, I recommend the authors restate any limitations in their model that they mentioned earlier in the text (e.g., 0.9v limitation) and provide an honest discussion regarding the limits of their model.
    {
        \color{blue}

    }
}

>>>>

\usepackage{cite}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{url}
\usepackage[usenames, dvipsnames]{color}

@@

@@

Electrodes in the laboratory are typically tested in a saline solution selected to mimic the circumstances in which they will operate when implanted. In an implanted setting there is no reference electrode and many of the well established electrochemical measurement regimes do not apply. Here we measure the behaviour an electrode/electrolyte situation as it would be seen from an implant device.

@@

    \begin{figure}
        \begin{center}
            \includegraphics[width=230pt]{graphics/interfaceSchematic_noMemristive}
        \end{center}
        \caption{Electrical schematic of the electrode-interface impedance model used in this work. $D_{a}$ and $D_{b}$ represent diodes; $CPE$ and $R_{s}$ represent the constant phase element and series resistance respectively.}
        \label{fig:schematic}
    \end{figure}

    In this manuscript we use an adapted version of the compact non-linear model presented in \cite{ScottSingle2013}. The model is compact in that it is suitable for entering into SPICE, i.e., freely available electrical circuit simulation software. The schematic for the model used here is presented in Figure \ref{fig:schematic}.
}

The electrical interface model of \cite{Franks2005} and \cite{ScottSingle2013} has two parts, a displacement branch and a Faradaic branch.
{
    \color{blue}
    Scott and Single incorporate memristors into their model, used to mimic a diffusion limited current conduction mode, i.e., their purpose is to throttle electrical current associated with Faradaic reactions. In an implanted setting an electrode should never be operated in a Faradaic conduction mode and certainly never to the point of becoming diffusion limited; for this reason, memristors have been omitted.
}

It has been suggested that the primary coupling mechanism between platinum electrodes and physiological saline is the reversible evolution of gas, probably oxygen, at the platinum surface, which dissolves without bubble formation.\cite{Greatbatch1969} Understanding the exact mechanism of displacement or Faradaic charge transfer at the electrode surface is beyond the scope of this work. Rather, we wish to characterise the electrode/electrolyte system in terms of a {\color{blue} small number of} parameters that may be fed into an electrical model.

@@

The electrical impedance between two electrodes in an electrolyte arises from two interface impedances in series with the resistance due to the bulk of the electrolyte itself. To quantify the impedance presented by a single interface we first need to determine inter-electrode resistance presented by the electrolyte bulk. {\color{blue} Finding these resistances allows us to quantify the interface's impedance via subtraction.} As there are always two interface impedances between any two electrodes, inter-electrode resistances cannot be measured in isolation.

In \cite{ScottSingle2013}, a series of transresistance measurements were used in conjunction with a physical description of the electrode geometry to build up a representative network of resistors. That resistor network was defined using five parameters: $R_{eri}$, $R_{lri}$, $R_{li}$, mesh depth, and edge depth. Readers are referred to \cite{ScottSingle2013} for the interpretation of these model parameters. Repeating that work, we have created a network of resistors that mimic the resistances due to the solution's bulk conductivity.

The mesh depth and padding values of five columns and three rows, respectively, has been taken directly from \cite{ScottSingle2013}. {\color{blue} These parameters, together with an array of 8 electrodes, leads to a generated network of 205 resistors, the layout of which is presented as figure \ref{fig:mesh}.}

A numerical fit is then made to the two independent scaling parameters $R_{eri}$ and $R_{li}$
\footnote{The values of $R_{eri}$ and $R_{sri}$ are defined by the lengths of the electrode and inter-electrode spacers respectively as described in \cite{ScottSingle2013}. The two parameters are proportional to each other by the ratio of their lengths. Hence, we define $R_{sri}$ to be 3/4 that of $R_{eri}$ making it a dependant parameter.}.
The resulting fit and parameter values are presented in figure~\ref{fig:transresistance} and table~\ref{tab:RESparams}.

Measurements of the inter-electrode impedance are made using a Tektronix TPS2014 DSO with four fully isolated and floating channels; an Agilent 33220A Function Generator; and a desktop PC {\color{blue} running Python 2.7 for instrumentation control and result processing.} The pH level of the stock solution is measured using a calibrated EDT Instruments BA-350 pH meter. It has a pH of 7.4 and no pH adjustments were made to the derived solutions.

{
    \color{blue}
    To be clear, the resistive mesh created in this step models \textit{only} the resistivity of the solution bulk. No contribution from the interface itself appears in these measurements. The is because voltages are always measured across pairs of non-driven electrodes and the measurements are of suitably high impedance. We determine the resistive component of the interface, $R_{S}$, in the next section.
}

@@

@@

\section{Displacement Parameters {\color{blue}/ CPE}}
\label{sect:cpe}
Displacement currents are responsible for capacitive behaviour at interfacial boundaries.
They are brought about by the redistribution of charge in response to applied fields, i.e. a charged electrode will attract or repel $Cl^{-}$ from its surface and reorientate polar molecules in the surrounding solution.\cite{Merrill05} This capacitive behaviour may also be the result of electrode polarisation in the form of reversible Faradaic reactions at the surface of the electrode.
Such reactions involving water and platinum, as identified in \cite{Horch2004,Mohtashami2011,Merrill05}, include:
\begin{eqnarray}
    Pt + H_{2}O &\Leftrightarrow& PtO + 2 H^{+} + 2 e^{-}\\
    PtO + H_{2}O &\Leftrightarrow& PtO_{2} + 2 H^{+} + 2e^{-}\\
    Pt + H^{+} & \Leftrightarrow & Pt-H\\
    Pt + H_{2}O + e^{-} &\Leftrightarrow& Pt-H+OH^{-}
\end{eqnarray}

Displacement currents are capable of transferring charge between electrical and biological systems without damaging electrodes or tissue.\cite{Horch2004}
This behaviour is modelled by way of a Constant Phase Element (CPE), also known as a fractional capacitor.
For simulation purposes this element is realised as an array of R-C branches.~\cite{ScottSingle2013,Morrison59,Elwakil10} where the value of each R-C pair is chosen to set the slope of impedance relative to frequency. The placement of R-C pairs within the CPE is determined by Morrison's parameters $m$ and $k$, where $k$ determines the spacing, or density, of R-C branches and $m$ determines the resulting slope. {\color{blue} The R-C pairs were generated via Python script to span 1\thinspace nHz--1\thinspace GHz with a spacing of 3 pairs per decade. The final CPE contains 55 branches i.e. 110 elements are used to create the CPE model.}

@@

\begin{figure}
    \begin{center}
        \includegraphics{graphics/chargeInjectionEffectiveVsFrequency_magnitude}
    \end{center}
    \caption{\color{blue} Per phase charge injected during CPE measurement in each concentrations of PBS, calculated using an effective surface area of 14\thinspace mm$^{2}$. {\color{red} IS THIS CHEATING? (I have given the charged injected in terms of effective surface area instead of geometric)}}
    \label{fig:chargeInjectionVsFrequency}
\end{figure}

{
    \color{blue}
    Current density during measurements, calculated using an effective surface area of $14\thinspace mm^{2}$, peaked at $161\thinspace \mu A/cm^{2}$ at 10\thinspace kHz in 1.0X PBS and fell as low as $197\thinspace nA/cm^{2}$ at 50\thinspace mHz in 0.025X PBS.
}Charge injection per phase is shown in figure~\ref{fig:chargeInjectionVsFrequency} for each solution.
The stimulus waveform was set at 300\thinspace mV (peak) at each measurement point. The measurement instruments used are the same as those used to measure the inter-electrode resistance.
Figure~\ref{fig:CPE_Magnitude} shows the magnitude of the measured response of each of the six PBS concentrations accompanied by simulation results. Figure~\ref{fig:CPE_Phase} shows the phase response for the same measurements, again with the simulated response at each concentration. The simulated responses are generated using the optimised parameter values presented in Table~\ref{tab:CPEparams}. Those values were chosen as they give the closest match to the measured data.

@@

Our model uses two reverse-connected diodes to represent Faradaic current conduction between electrode and electrolyte. Specifically we use the diodes to model irreversible Faradaic reactions. (Any reversible reactions, those producing reactants bound to the surface of the electrode, are encapsulated by the displacement branch (CPE) of the model.)
Possible irreversible Faradaic reactions for platinum electrodes {\color{blue}(each with a different reaction potential)} in saline are:

@@

{
    \color{blue}
We use a diode to model the electrical current conduction caused by one reaction proceeding in one direction, i.e. a diode is the circuit representation of the Butler-Volmer equation for a single reaction proceeding in a single direction. 
    We wish to model the onset of Faradaic current as a way of knowing that an electrode has been pushed ``too far''. It is expected that the situation in-vivo will be much more complex and that fitted diode parameters may encapsulate multiple reactions if they occur close together.
}

We measure Faradaic conduction by separating it from the displacement behaviour of the CPE. By using a DC measurement technique, as opposed to cyclic measurements, we are able to separate the effects in time.
Like a conventional capacitor, when the voltage across the CPE changes, it responds by absorbing or ejecting charge, causing a spike in current. The CPE will draw negligible current once it has settled and in the case of our model we assume the remaining current conduction to be the result of Faradaic processes.

The equation governing current conduction in a diode where the forward voltage across the diode, $V_{D}$, is small is
\begin{equation}
    I = i_{0}  e^{V_{D} / n V_{T}}
\end{equation}
where $V_{T}$ is the thermal voltage and is approximately 25\thinspace mV at a temperature of 300\thinspace K. We wish to find how the saturation current ($i_{0}$) and the ideality factor ($n$) scale as the solution salinity is varied.

@@

In \cite{Greatbatch1969} it is shown that the stirring of an air-saturated solution of physiological saline reduces the settling time of platinum electrodes to stimulus transients; with both situations eventually settling to the same impedance.

@@

@@

\begin{figure}
    \begin{center}
        \includegraphics{graphics/sheep_transimpedance_doubleFit_mag}
    \end{center}
    \caption{\color{blue} Comparison between the magnitude of in-vivo transresistance measurements (blue dots) and simulation results using fitted mesh parameters (red trace).}
    \label{fig:transimpedance_sheep_mag}
\end{figure}

\begin{figure}
    \begin{center}
        \includegraphics{graphics/sheep_transimpedance_doubleFit_phase}
    \end{center}
    \caption{\color{blue} Comparison between the phase of in-vivo transresistance measurements (blue dots) and simulation results using fitted mesh parameters (red trace). The simulation mesh does not capture any phase relationship due to being purely resistive.}
    \label{fig:transimpedance_sheep_phase}
\end{figure}

@@

{\color{blue}
    Four measurements of displacement current (CPE element in the model) were} taken at regular intervals over a period of 30 minutes and have been presented in figures~\ref{fig:displacement_sheepCPEMagnitude} and~\ref{fig:displacement_sheepCPEPhase}.
{\color{blue} These measurements appear on top of the traces taken from figures~\ref{fig:CPE_Magnitude} and~\ref{fig:CPE_Phase} for comparison.}

@@

\begin{figure}
    \begin{center}
        \includegraphics{graphics/displacement-withSheep_impedanceVsFrequency_phase}
    \end{center}
    \caption{{\color{blue} Comparison between the mean phase response of measured displacement currents in four live sheep (blue trace), each of the PBS traces taken from figure~\ref{fig:CPE_Phase} (grey lines), and a simulated fit to the sheep data (red trace). Error bars have been placed one standard deviation either side of the mean for measured data.}}
    \label{fig:displacement_sheepCPEPhase}
\end{figure}

\section{Conclusion}
\label{sect:conclusion}
We have used the parameters of a compact electrical model of an implantable electrode as a way of objectively comparing scenarios.
We have measured how changing the concentration of PBS affects the parameters of this model.
We have drawn direct relationships with concentration {\color{blue} and conductivity} sufficient to predict interface characteristics at arbitrary dilutions of PBS.
We found that the magnitude of the CPE element moves much more slowly with concentration whereas the resistance moved almost linearly.
Using DC measurements we conclude that the {\color{blue} concentration of PBS has no effect on the onset of Faradaic conduction for the range of concentrations investigated.} We hypothesise that this is because the double-layer sets the effective concentration of ions{\color{blue}, those involved in the Faradaic reactions,} in a localised volume surrounding each electrode.
In-vivo measurements of platinum electrodes in sheep spine show that no single concentration of PBS matches both the CPE and the resistive characteristics at once. {\color{blue} These measurements also reveal a considerable amount of reactance (see transimpedance measurements in-vivo) that is captured neither by PBS nor the model presented when using a purely resistive spreading resistance.} Nevertheless, {\color{blue}if a single solution of PBS is to be used as a human phantom, a one-tenth concentration solution (by mass)} is a good compromise.

{
    \color{blue}
    While the model used is limited to predicting only the onset of Faradaic current conduction, it usefully captures the transition into this biologically destructive stimulus regime sufficient for it to be avoided.

    The final model, including 205 lines of resistive mesh, is represented by a SPICE netlist comprised of 348 lines. Simulation of this model, using the freely available ngSpice circuit simulator, at 1000 frequencies between spaced between 50\thinspace mHz to 10\thinspace kHz takes approximately 1 second on a modern CPU. 
}
{
    \color{red}
    How is this last line? It's new information in the conclusion...
}

\section*{Acknowledgment}
The authors wish to acknowledge the The University of Waikato for financial support and Saluda Medical for the use of facilities and equipment. {\color{red}Special thanks are expressed to Peter Single of Saluda Medical for help conducting in-vivo measurements.}

@@

\bibitem{ScottSingle2013}
    Jonathan Scott and Peter Single,
    ``Compact Nonlinear Model of an Implantable Electrode Array for Spinal Cord Stimulation (SCS)'',
    to be published in
    {\em IEEE Transactions on Biomedical Circuits and Systems},
    DOI: 10.1109/TBCAS.2013.2270179, 2013.

@@

    {
        \color{blue}
        The diodes are there to model the change in current attributed to the first biologically damaging reaction that occurs at the electrode/electrolyte interface. Whatever reaction occurs first in the relevant electrolyte is the reaction that diode aims to mimic. We observe that the current increases near exponentially around this point so we use a diode to mimic that. We expect this reaction may not be the same in-vivo as it is in-vivo but the properties of a diode can be altered to match the situation.

        For secondary reactions occurring at higher potentials, a second pair of diodes with the properties of that specific reaction can be added. The problem here is that memristors will be needed (as used in ScottSingle2013) to account for reactant depletion. We don't account for reactant depletion for the same reason we don't account for reactions other than the onset of the first Faradaic reaction(s) taking place. The reason is that we never want biologically damaging reactions to occur due to an implant, so if they happen then something has gone wrong.

        Think of our model of Faradaic reactions as a ``you've pushed this too far'' indicator for an electrical engineer.

        We have added a statement to address this concern directly after equations 5--9 in section \ref{sect:faradaic}.
    }

@@

    {
        \color{blue}
        That is correct, we are not aiming to determine what is happening at the electrode surface. We have a flexible model that can be used to simulate interfacial impedance and we use this model to compare this impedance for electrodes in various electrolytes. If the solution changes, the reactions change. Since we are only interested in the onset of Faradaic conduction, and we expect this to change when we move to different solutions, we do not want to rely on the exact electrochemistry at play.

@@

    {
        \color{blue}
        This is a good point. ScottSingle2013 compare the models prediction to stimulus waveforms. Our aim is to capture how the model parameters change as the concentration of PBS is varied and to compare those changes to parameters obtained from in-vivo measurements using a live sheep. Were doing this because 0.1X PBS is commonly used as a phantom for real spine. There hasn't been an objective way of determining how accurate that assumption was until we used the model to make this comparison. In future we hope to match it against in-vivo stimulus waveforms; at present it is out of scrope.
    }

@@

    {
        \color{blue}
        We need this because it represents the resistance of the interface itself, separate from that of the fluid.
        We measure the impedance magnitude of the displacement currents at high frequency (flattened part of Fig. 5). For each concentration the impedance magnitude at 10kHz is higher than is predicted by the transimpedance measurements made in Fig. 3. The transimpedance measurements do not measure impedance across an interface that is passing current and therefore any voltage drop across an interface due to $R_{S}$ vanishes. The displacement measurements take the voltage across interfaces that are passing stimulus currents. These measurements see the potential drop across the interface due to the current flow (because of Ohms law).

@@

    {
        \color{blue}
        That is interesting and definitely important to our modelling work. We had presented the charge injected in terms of the geometric surface area of each electrode which we know is less than the real (or effective) surface area, estimated to be $14\thinspace mm^{2}$. Calculating the charge injection using the effective surface area is more relevant here and gives results consistant with Brummer \& Roblee (Annals New York Academy of Sciences, 159-171, 1983), who also express their values in terms of ``real'' area.

    }
    {
        \color{red}
        How does this read now? In your last comment you said...

        We have to acknowledge that this is anomalous. We were stirring, but definitely observed no gas. Say just as you say, the results are consistent with it NOT being a problem, although other researchers have observed otherwise. We seem to be saying that they are wrong, so we should say just that, we observe something different, yet results are clearly consistent and explicable!
    }

@@

    {
        \color{blue}
        The RC pairs are generated via a Python script for a given frequency range. The frequency range supplied for all simulations in this manuscript was 1.0 nHz to 1.0 GHz with R-C pairs spaced 3 per decade. Adding more pairs than necessary came at negligible computation cost and having the circuit generated in an automated fashion adds no extra work.
        {
            \color{red}
            We would also like to note that the effect of the CPE is far from that of a simple capacitor, i.e. far from being exponential. This is part of the reason they are so difficult to simulate in a compact circuit simulator.
        }
    }

@@

    {
        \color{blue}
        Yes you are quite right. We think the results of the CPE in-vivo vs. PBS are intriguing in light of what the previous literature would have you expect. We also know of no satisfactory explanation of what the CPE actually is. One of the authors came up against this uncertainty while writing \cite{ScottSingle2013}.

        We feel that the in-vivo results will be affected by the significant inhomogeneity of real spine by proximity of bone, change in fluid at dura, etc. Perhaps you may know of a way to clarify this issue?
    }

    {
       \color{red}
       Is this what you were after? 
    }

@@

    9. In their conclusions, I recommend the authors restate any limitations in their model that they mentioned earlier in the text (e.g., 0.9v limitation) and provide an honest discussion regarding the limits of their model.
    {
        \color{blue}
        The 0.9V limitation is rather arbitrary as it was specific to the use of electrodes 2 and 7. We have instead clarified that the model only captures the onset of Faradaic conduction sufficient for it to be avoided. A sentence to this effect have been added to the end of the conclusion. 
    }
}
